---
title: Intro to Machine Learning 
subtitle: Airline Satisfaction Classification
author: Michael Crosson, Saaket Joshi, Luke Leon, Quinlan O'Connell, Austin Yeh
date: 07/26/2024
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ISLR2)
library(pROC)
library(glmnet)
library(tidyverse)
library(reshape2)
library(caret)
library(randomForest)
library(tree)
library(rpart)
library(rpart.plot)
library(BART)
library(patchwork)
library(corrplot)
library(gbm)

set.seed(120)
```

# Import Dataset
```{r}
trainData <- read_rds("data/airline_encoded_train.rds")
testData <- read_rds("data/airline_encoded_test.rds")

df <- rbind(trainData, testData)
```

# Exploratory Data Analysis (EDA)
```{r}
df %>% glimpse
```

## Change DataTypes
```{r}
df <- df %>% mutate(
  Inflight.wifi.service = as.factor(Inflight.wifi.service),
  Departure.Arrival.time.convenient = as.factor(Departure.Arrival.time.convenient),
  Ease.of.Online.booking = as.factor(Ease.of.Online.booking),
  Gate.location = as.factor(Gate.location),
  Food.and.drink = as.factor(Food.and.drink),
  Online.boarding = as.factor(Online.boarding),
  Seat.comfort = as.factor(Seat.comfort),
  Inflight.entertainment = as.factor(Inflight.entertainment),
  On.board.service = as.factor(On.board.service),
  Leg.room.service = as.factor(Leg.room.service),
  Baggage.handling = as.factor(Baggage.handling),
  Checkin.service = as.factor(Checkin.service),
  Inflight.service = as.factor(Inflight.service),
  Cleanliness = as.factor(Cleanliness)
)
```


## Summary
```{r}
summary(df)
```
## Univariate

```{r}
p1 <- ggplot(df, aes(Age)) + 
  geom_density(fill="lightblue", color="black", alpha=0.8) +
  labs(title = "Distribution of Age") + theme_bw()

p2 <- df %>%
  mutate(satisfaction = recode(satisfaction, `0` = "neutral or dissatisfied", `1` = "satisfied")) %>%
  ggplot(aes(x = Age, fill = satisfaction)) +
  geom_density(alpha = 0.5) +
  labs(x = NULL) +
  theme_bw() +
  facet_wrap(~satisfaction)

p1/p2
```

Most people who booked flights are around 25 years old and 40 years old.


```{r}
p1 <- ggplot(df, aes(Flight.Distance)) + 
  geom_density(fill = "lightblue", color = "black", alpha = 0.8) +
  labs(title = "Distribution of Flight Distance") + 
  theme_bw()

p2 <- df %>%
  mutate(satisfaction = recode(satisfaction, `0` = "neutral or dissatisfied", `1` = "satisfied")) %>%
  ggplot(aes(x = Flight.Distance, fill = satisfaction)) +
  geom_density(alpha = 0.5) +
  labs(x = NULL, title = "Distribution of Flight Distance by Satisfaction") +
  theme_bw() +
  facet_wrap(~ satisfaction)

combined_plot <- p1 / p2

print(combined_plot)
```

Majority of flights are within 1500 miles

```{r}
p1 <- ggplot(df, aes(Departure.Delay.in.Minutes)) + 
  geom_density(fill = "lightblue", color = "black", alpha = 0.8) +
  labs(title = "Distribution of Departure Delay in Minutes") + 
  theme_bw()

p2 <- df %>%
  mutate(satisfaction = recode(satisfaction, `0` = "neutral or dissatisfied", `1` = "satisfied")) %>%
  ggplot(aes(x = Departure.Delay.in.Minutes, fill = satisfaction)) +
  geom_density(alpha = 0.5) +
  labs(x = NULL, title = "Distribution of Departure Delay by Satisfaction") +
  theme_bw() +
  facet_wrap(~ satisfaction)

combined_plot <- p1 / p2

print(combined_plot)
```


```{r}
df %>% 
  select(8:15) %>%
  gather(variable, value) %>%
  ggplot(aes(x = value)) +
  geom_bar(fill = "salmon") +
  facet_wrap(~ variable, scales = "free_x") +
  theme_bw() +
  labs(title = "Range Variables pt.1")

df %>% 
  select(16:21) %>%
  gather(variable, value) %>%
  ggplot(aes(x = value)) +
  geom_bar(fill = "salmon") +
  facet_wrap(~ variable, scales = "free_x") +
  theme_bw() +
  labs(title = "Range Variables pt.2")

```




## Correlation Matrix (pt.1)

[Reference to the correlation matrix code](https://towardsdatascience.com/how-to-create-a-correlation-matrix-with-too-many-variables-309cc0c0a57#:~:text=With%20more%20variables%2C%20it%20may,results%20are%20readable%20and%20useful.)

```{r}
corr_simple <- function(data=df,sig=0.45){
  #convert data to numeric in order to run correlations
  #convert to factor first to keep the integrity of the data - each value will become a number rather than turn into NA
  df_cor <- data %>% mutate_if(is.character, as.factor)
  df_cor <- df_cor %>% mutate_if(is.factor, as.numeric)
  #run a correlation and drop the insignificant ones
  corr <- cor(df_cor)
  #prepare to drop duplicates and correlations of 1     
  corr[lower.tri(corr,diag=TRUE)] <- NA 
  #drop perfect correlations
  corr[corr == 1] <- NA 
  #turn into a 3-column table
  corr <- as.data.frame(as.table(corr))
  #remove the NA values from above 
  corr <- na.omit(corr) 
  #select significant values  
  corr <- subset(corr, abs(Freq) > sig) 
  #sort by highest correlation
  corr <- corr[order(-abs(corr$Freq)),] 
  #print table
  print(corr)
  #turn corr back into matrix in order to plot with corrplot
  mtx_corr <- reshape2::acast(corr, Var1~Var2, value.var="Freq")
  #plot correlations visually
  corrplot(mtx_corr, is.corr=FALSE, tl.col="black", na.label=" ", tl.srt=45, cl.ratio = 0.2)
}
corr_simple(data = df %>% select(-GenderMale, -Loyal.Customer, -Business.Travel, -ClassBusiness, -ClassEco.Plus), sig=.45)
```

## Multivariate between Predictors

```{r}
ggplot(df, aes(x = Ease.of.Online.booking, y = Inflight.wifi.service)) +
  geom_bin2d() +
  scale_fill_gradient(low = "white", high = "blue")
```
```{r}
ggplot(df, aes(x = Inflight.entertainment, y = Cleanliness)) +
  geom_bin2d() +
  scale_fill_gradient(low = "white", high = "blue")
```

```{r}
viz_df <- df %>%
  mutate(distance_bin = cut(Flight.Distance, breaks = 70, labels = FALSE),
         Seat.comfort = as.numeric(Seat.comfort))

summary_df <- viz_df %>%
  group_by(distance_bin) %>%
  summarise(
    mean_seat_comfort = mean(Seat.comfort),
    bin_center = mean(Flight.Distance)
  )

ggplot(summary_df, aes(x = bin_center, y = mean_seat_comfort)) +
  geom_line() +
  geom_smooth() +
  labs(
    title = "Flight Distance vs Seat Comfort",
    x = "Flight Distance (binned)",
    y = "Mean Seat Comfort"
  )
```


```{r}
viz_df <- df %>%
  mutate(distance_bin = cut(Flight.Distance, breaks = 70, labels = FALSE),
         On.board.service = as.numeric(On.board.service))

summary_df <- viz_df %>%
  group_by(distance_bin) %>%
  summarise(
    mean_onboard_service = mean(On.board.service),
    bin_center = mean(Flight.Distance)
  )

ggplot(summary_df, aes(x = bin_center, y = mean_onboard_service)) +
  geom_line() +
  geom_smooth() +
  labs(
    title = "Flight Distance vs Onboard Service",
    x = "Flight Distance (binned)",
    y = "Mean Onboard Service"
  )
```


```{r}
viz_df <- df %>%
  mutate(age_bin = cut(Age, breaks = 70, labels = FALSE),
         On.board.service = as.numeric(On.board.service))

summary_df <- viz_df %>%
  group_by(age_bin) %>%
  summarise(
    mean_onboard_service = mean(On.board.service),
    bin_center = mean(Age)
  )

ggplot(summary_df, aes(x = bin_center, y = mean_onboard_service)) +
  geom_line() +
  geom_smooth() +
  labs(
    title = "Age vs Onboard Service",
    x = "Age (binned)",
    y = "Onboard Service"
  )
```

```{r}
viz_df <- df %>%
  mutate(age_bin = cut(Age, breaks = 50, labels = FALSE),
         Baggage.handling = as.numeric(Baggage.handling))

summary_df <- viz_df %>%
  group_by(age_bin) %>%
  summarise(
    mean_bag_hand = mean(Baggage.handling),
    bin_center = mean(Age)
  )

ggplot(summary_df, aes(x = bin_center, y = mean_bag_hand)) +
  geom_line() +
  geom_smooth() +
  labs(
    title = "Age vs Baggage Handling",
    x = "Age (binned)",
    y = "Mean Baggage Handling"
  )
```


```{r}
viz_df <- df %>%
  mutate(age_bin = cut(Age, breaks = 70, labels = FALSE),
         Ease.of.Online.booking = as.numeric(Ease.of.Online.booking))

summary_df <- viz_df %>%
  group_by(age_bin) %>%
  summarise(
    mean_ease = mean(Ease.of.Online.booking),
    bin_center = mean(Age)
  )

ggplot(summary_df, aes(x = bin_center, y = mean_ease)) +
  geom_line() +
  geom_smooth() +
  labs(
    title = "Age vs Ease of Online Booking",
    x = "Age (binned)",
    y = "Mean Ease of Online Booking"
  )
```




# Models

## Setup
```{r}
conf_mats <- c()
```


## K-Nearest-Neighbors Classification
```{r}
# Define the train control with 10-fold cross-validation
train_control <- trainControl(method = "cv", number = 10)

set.seed(120)
knn_model <- train(
  satisfaction ~ ., data = trainData,
  method = "knn",
  trControl = train_control,
  metric = "Accuracy",
  tuneLength = 5
)
```

### ROC
```{r}
predictions <- predict(knn_model, newdata = testData, type = "prob")[,2]
predicted_classes <- ifelse(predictions > 0.5, 1, 0)

# Create ROC curve
roc_curve <- roc(testData$satisfaction, predictions)
plot(roc_curve, main = "ROC Curve for KNN Model")

# Calculate and display AUC
auc_value <- auc(roc_curve)
text(0.6, 0.2, paste("AUC =", round(auc_value, 4)), col = "blue")
```

### Confusion Matrix
```{r}
conf_matrix <- confusionMatrix(as.factor(predicted_classes), as.factor(testData$satisfaction))
conf_mats <- c(conf_mats, c("knn_model" = conf_matrix$overall['Accuracy']))
conf_matrix
```


## Logistic Regression w/ Lasso

```{r}
set.seed(120)
# Train the LASSO logistic regression model
lasso_model <- train(satisfaction ~ ., 
                     data = trainData, 
                     method = "glmnet", 
                     family = "binomial", 
                     trControl = train_control,
                     tuneLength = 10)
```

### Tune Parameters
```{r}
# Display the best model
lasso_model$bestTune
```

### ROC
```{r}
predictions <- predict(lasso_model, newdata = testData, type = "prob")[,2]
predicted_classes <- ifelse(predictions > 0.5, 1, 0)

roc_curve <- roc(testData$satisfaction, predictions)
plot(roc_curve, main="ROC Curve for LASSO Logistic Regression Model")

auc <- auc(roc_curve)
text(0.6, 0.2, paste("AUC =", round(auc, 4)), col = "blue")
```

### Confusion Matrix
```{r}
conf_matrix <- confusionMatrix(as.factor(predicted_classes), as.factor(testData$satisfaction))
conf_mats <- c(conf_mats, c("lasso_model" = conf_matrix$overall['Accuracy']))
conf_matrix
```




## Regular Decision Tree

```{r}
set.seed(120)
tree_model <- rpart(satisfaction ~ ., data = trainData, method = "class")
```

### Tree Model
```{r}
rpart.plot(tree_model, type = 3, extra = 102, fallen.leaves = TRUE, 
           main = "Decision Tree for Satisfaction", 
           box.palette = "gray", shadow.col = "gray", nn = TRUE)
```

### ROC
```{r}
predictions <- predict(tree_model, newdata = testData, type = "prob")[,2]
predicted_classes <- ifelse(predictions > 0.5, 1, 0)

# Create ROC curve
roc_curve <- roc(testData$satisfaction, predictions)
plot(roc_curve, main = "ROC Curve for Single Tree Model")

# Calculate and display AUC
auc_value <- auc(roc_curve)
text(0.6, 0.2, paste("AUC =", round(auc_value, 4)), col = "blue")
```

### Confusion Matrix
```{r}
conf_matrix <- confusionMatrix(as.factor(predicted_classes), as.factor(testData$satisfaction))
conf_mats <- c(conf_mats, c("tree_model" = conf_matrix$overall['Accuracy']))
conf_matrix
```

## Bagging 
```{r}
set.seed(120)
bag_fit <- train(satisfaction ~., data = trainData, 
                 method = "treebag",
                 trControl = train_control)
```

### ROC
```{r}
predictions <- predict(bag_fit, newdata = testData, type = "prob")[,2]
predicted_classes <- ifelse(predictions > 0.5, 1, 0)

# Create ROC curve
roc_curve <- roc(testData$satisfaction, predictions)
plot(roc_curve, main = "ROC Curve for Bagging Model")

# Calculate and display AUC
auc_value <- auc(roc_curve)
text(0.6, 0.2, paste("AUC =", round(auc_value, 4)), col = "blue")
```

### Confusion Matrix
```{r}
conf_matrix <- confusionMatrix(as.factor(predicted_classes), as.factor(testData$satisfaction))
conf_mats <- c(conf_mats, c("bag_fit" = conf_matrix$overall['Accuracy']))
conf_matrix
```

## Boosting
```{r include=FALSE}
trainData$satisfaction <- as.numeric(trainData$satisfaction) - 1
testData$satisfaction <- as.numeric(testData$satisfaction) - 1
```


```{r}
set.seed(120)
# Train the Gradient Boosting Model
gbmfit <- gbm(satisfaction ~ ., 
              data = trainData, 
              distribution = "bernoulli",
              cv.folds = 10)
```

### ROC
```{r}
predictions <- predict(gbmfit, newdata = testData, type = "response")
predicted_classes <- ifelse(predictions > 0.5, 1, 0)

# Create ROC curve
roc_curve <- roc(testData$satisfaction, predictions)
plot(roc_curve, main = "ROC Curve for Boosting Model")

# Calculate and display AUC
auc_value <- auc(roc_curve)
text(0.6, 0.2, paste("AUC =", round(auc_value, 4)), col = "blue")
```



### Confusion Matrix
```{r}
conf_matrix <- confusionMatrix(as.factor(predicted_classes), as.factor(testData$satisfaction))
conf_mats <- c(conf_mats, c("gbmfit" = conf_matrix$overall['Accuracy']))
conf_matrix
```

```{r include=FALSE}
trainData$satisfaction <- as.factor(trainData$satisfaction)
testData$satisfaction <- as.factor(testData$satisfaction)
```


## BART
```{r}
trainData[,c("GenderMale", "Loyal.Customer", "Business.Travel", "ClassBusiness", "ClassEco.Plus", "satisfaction")] <- trainData[,c("GenderMale", "Loyal.Customer", "Business.Travel", "ClassBusiness", "ClassEco.Plus", "satisfaction")] %>% sapply(as.numeric) - 1

trainData[,c("GenderMale", "Loyal.Customer", "Business.Travel", "ClassBusiness", "ClassEco.Plus", "satisfaction")] <- trainData[,c("GenderMale", "Loyal.Customer", "Business.Travel", "ClassBusiness", "ClassEco.Plus", "satisfaction")] %>% sapply(as.logical)

testData[,c("GenderMale", "Loyal.Customer", "Business.Travel", "ClassBusiness", "ClassEco.Plus", "satisfaction")] <- testData[,c("GenderMale", "Loyal.Customer", "Business.Travel", "ClassBusiness", "ClassEco.Plus", "satisfaction")] %>% sapply(as.numeric) - 1

testData[,c("GenderMale", "Loyal.Customer", "Business.Travel", "ClassBusiness", "ClassEco.Plus", "satisfaction")] <- testData[,c("GenderMale", "Loyal.Customer", "Business.Travel", "ClassBusiness", "ClassEco.Plus", "satisfaction")] %>% sapply(as.logical)
```


```{r}
xtrain <- trainData %>% select(-satisfaction) %>% as.data.frame()
ytrain <- trainData$satisfaction

xtest <- testData %>% select(-satisfaction) %>% as.data.frame()
ytest <- testData$satisfaction

set.seed(120)
bart_model <- lbart(xtrain, ytrain, x.test = xtest, ntree = 50, ndpost = 1000, nskip = 100)
```

### ROC
```{r}
predictions <- bart_model$prob.test.mean
predicted_classes <- ifelse(predictions > 0.5, 1, 0)

# Create ROC curve
roc_curve <- roc(testData$satisfaction, predictions)
plot(roc_curve, main = "ROC Curve for BART Model")

# Calculate and display AUC
auc_value <- auc(roc_curve)
text(0.6, 0.2, paste("AUC =", round(auc_value, 4)), col = "blue")
```

### Confusion Matrix
```{r}
test <- factor(testData$satisfaction, levels = c(FALSE, TRUE))
levels(test) <- c(0, 1)

conf_matrix <- confusionMatrix(as.factor(predicted_classes), test)
conf_mats <- c(conf_mats, c("bart_model" = conf_matrix$overall['Accuracy']))
conf_matrix
```


# Model Comparison
```{r}
accuracy_df <- 
  data.frame(conf_mats, 
             Model = names(conf_mats), 
             Accuracy = as.numeric(conf_mats)) %>% 
  select(-conf_mats) %>% 
  mutate(Model = gsub("\\.Accuracy", "", Model)) %>% 
  mutate(Model = reorder(Model, -Accuracy))

rownames(accuracy_df) <- NULL

accuracy_df
```

```{r}

ggplot(accuracy_df, aes(x = Model, y = Accuracy)) +
     geom_bar(stat = "identity", fill = "steelblue") +
     geom_text(aes(label = sprintf("%.2f%%", Accuracy * 100)), vjust = -0.3) +
     labs(
         title = "Model Accuracy Comparison",
         x = "Model",
         y = "Accuracy"
     ) +
     theme(axis.text.x = element_text(angle = 45, hjust = 1))
```




